Making Higher-Order Superposition Work

### 2022-03-16 scratch/paper.tex
hle

\newlength{\myskip}
\setlength{\myskip}{0.6ex}
% TYPESETTING: when two environments are next to
% each other, we need to somehow squeeze them
\newcommand{\removemyskip}{\vspace{-\myskip}}

\newenvironment{theoremx}% environment name
{% begin code
  \vspace{\myskip}
  \begin{theorem}
}%
{\end{theorem}
 \vspace{\myskip}}% end code

\newenvironment{examplex}% environment name
{% begin code
  \vspace{\myskip}
  \begin{example}
}%
{\end{example}
 \vspace{\myskip}}% end code

@@

\newenvironment{definitionx}% environment name
 {% begin code
   \vspace{\myskip}
   \begin{definition}
 }%
 {\end{definition}
  \vspace{\myskip}}% end code

@@

\begin{definitionx}
   Given a propositional literal $L$ and a propositional clause set~$N$, the
   set of \emph{propositional hidden literals} for $L$ and~$N$ is
   $\PHL(L, N) = \{ L' \mid L' \bigpimpl{N}^\ast L \} \setminus \{ L \}$, where $\bigpimpl{N}$ is
   defined such that $\neglit{L_1} \bigpimpl{N} L_2$ whenever $L_1 \cor L_2 \in N.$
   Moreover, $\PHL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n
   \PHL(L_i, N)$.
\end{definitionx}

@@

\begin{definitionx}
   \label{def:hle}\label{def:hte}
   Given $L' \cor L \cor C \in N,$ if $L' \in
   \PHL(L, N)$, \emph{hidden literal elimination} (HLE) replaces $N$ by $(N
   \setminus \{L' \cor L \cor C\}) \cup \{L \cor C\}$.
   Given $C \in N$, $\{L_1,\dots,L_n\} = \PHL(C, N)$, and $C' = C \cor L_1 \cor
   \cdots \cor L_n$, if $C'$ is a tautology, \emph{hidden
   tautology elimination} (HTE) replaces $N$ by $N \setminus \{C\}$.
\end{definitionx}

\begin{theoremx}
   The result of applying HLE or HTE to a clause set $N$ is
   equivalent to $N$.
\end{theoremx}

@@

\begin{definitionx}
\label{def:hl-fo}
Given a literal $L$ and a clause set $N$,
the set of \emph{hidden literals} for $L$ and $N$
is $\HL(L,N) = \{ L' \mid L' \bigfimpl{N}^\ast L \} \setminus \{ L \}$, where
$\bigfimpl{N}$ is defined so that
(1)~$\neglit{L'\sigma} \bigfimpl{N} L\sigma$ if
$L' \cor L \in N$ and $\sigma$ is a substitution;
(2)~$\eqlit{s}{t} \bigfimpl{N} \eqlit{u[s]}{u[t]}$ for all terms $s, t$ and
contexts $u[\phantom{\cdot}]$; and
(3)~$\neqlit{u[s]}{u[t]} \bigfimpl{N} \neqlit{s}{t}$ for all terms $s, t$ and
contexts $u[\phantom{\cdot}]$.
Moreover, $\HL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n \HL(L_i, N)$.
\end{definitionx}

@@

\begin{examplex}
   \label{example:hle}
   Consider the clause set $N = \{ \cst{p}(x) \cor \neglit{\cst{p}(\cst{f}(x))}{,}\;
   \cst{p}(\cst{f}(\cst{f}(x))) \cor \eqlit{\cst{a}}{\cst{b}} \}$ and the clause
   $C = \neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \cor \cst{p}(x)$. The first
   clause in $N$ induces $\cst{p}(\cst{f}(x)) \bigfimpl{N} {\cst{p}(x)}$,
   $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N} \cst{p}(\cst{f}(x))$, and
   hence $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N}^\ast  \cst{p}(x)$.
   Together with the second clause in $N$, it can be used to derive $\neqlit{\cst{a}}{\cst{b}}
   \bigfimpl{N}^\ast  \cst{p}(x)$. Finally, using rule~(3) of Definition
   \ref{def:hl-fo}, we derive $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})}
   \bigfimpl{N}^\ast  \cst{p}(x)$---that is,
   $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \in \HL(\cst{p}(x), N)$. This
   allows us to remove $C$'s first literal using \infname{HLE}.
\end{examplex}

@@

\begin{examplex}
   \label{example:unithle}
   \looseness=-1
   Consider the clause set $N = \{\cst{p}(x) \cor \cst{q}(\cst{f}(x)){,}\;
   \neglit{\cst{q}(\cst{f}(\cst{a}))} \cor \eqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}{,}\;
   \neqlit{\cst{f}(x)}{\cst{g}(y)}\}$ and the clause $C = \neglit{\cst{p}(\cst{a})} \cor
   \neglit{\cst{q}(\cst{b})}$. The first clause in $N$ induces
   $\neglit{\cst{q}(\cst{f}(\cst{a}))} \bigfimpl{N} \cst{p}(\cst{a})$, whereas
   the second one induces $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})} \bigfimpl{N}
   \neglit{\cst{q}(\cst{f}(\cst{a}))}$. Thus, we have $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}
   \bigfimpl{N}^\ast \cst{p}(\cst{a})$---that is, $\neqlit{\cst{f}(\cst{b})}{\cst{f}(\cst{c})} \in
   \HL(\cst{p}(\cst{a}), N)$. By applying the substitution $\{x \mapsto \cst{b}{,}\allowbreak\; y
   \mapsto \cst{c}\}$ to the third clause in $N$, we can fulfill the conditions of
   \infname{UnitHLE} and remove $C$'s first literal.
\end{examplex}

@@

\begin{definitionx}
   A clause $C$ is a \emph{hidden tautology} for a clause set $N$ if there
   exists a finite set $\{L_1,\dots,L_n\} \subseteq \HL(C,N)$ such that
   $C \cor L_1 \cor \cdots \cor L_n$ is a tautology.
\end{definitionx}
\removemyskip
\begin{examplex}
   \label{example:hidden-tautologies-completeness}
   \looseness=-1
   In general, hidden tautologies are not redundant and cannot be deleted during saturation.
   Consider the unsatisfiable set $N = \{ \negpredlit{\cst{a}}{,}\;
   \negpredlit{\cst{b}}{,}\; \pospredlit{\cst{a}} \cor
   \pospredlit{\cst{c}}{,}\allowbreak\; \pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}
   \}$, the order $\cst{a} \prec
   \cst{b} \prec \cst{c}$, and the empty selection function. The
   only possible superposition inference from $N$ is between the last two clauses,
   yielding the hidden tautology $\pospredlit{\cst{a}} \cor \pospredlit{\cst{b}}$
   (after simplifying away $\neqlit{\top}{\top}$), which is entailed by the larger
   clauses $\pospredlit{\cst{a}} \cor \pospredlit{\cst{c}}$ and
   $\pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}$. If this clause is removed,
   the prover could enter an infinite loop, forever generating and deleting the
   hidden tautology\confrep{}{ and never getting the opportunity to derive the
   empty clause}.
\end{examplex}

@@

\begin{theoremx}
The rules \infname{HLE}, \infname{FLE}, \infname{CongHLE$^+$},
\infname{Cong\allowbreak HLE$^-$}, \infname{UnitHLE}, \infname{HTR},
\infname{FLR}, and \infname{UnitHTR} are sound simplification rules.
\end{theoremx}

@@

\begin{definitionx}
   A predicate symbol is called \emph{singular}\confrep{ (or ``non-self-referential'')}{}
   for a clause set $N$ if it occurs at most once in every clause contained in
   $N.$
\end{definitionx}
\removemyskip
\begin{definitionx}
   \label{def:flat-res}
   Let $C = \pospredlit{\cst{p}(\tuplen{s})} \cor C'$ and $D =
   \negpredlit{\cst{p}(\tuplen{t})} \cor D'$ be clauses with no variables in
   common. The clause $\neqlit{s_1}{t_1} \cor \cdots \cor \neqlit{s_n}{t_n}
   \cor C' \cor D'$ is a \emph{flat resolvent} of $C$ and $D$ on $\cst{p}$.
\end{definitionx}

@@

\begin{definitionx}
\label{def:flat-res-set}
   Let $M, N$ be clause sets and $\cst{p}$ be a singular predicate for
   $M$. Let $\flatresiter$ be the following relation on clause set pairs and clause sets:
   \begin{enumerate}
   \item $(M{,}\> \{\arbpredlit{\cst{p}(\tupleempty{s})} \cor C'\} \uplus N) \flatresiter (M{,}\; N' \cup N)$ if
   $N'$ is the set that consists of all clauses (up to variable renaming) that are flat resolvents
   with $\arbpredlit{\cst{p}(\tupleempty{s})} \cor C'$ on $\cst{p}$ and a clause from $M$ as premises.
   The premises' variables are renamed apart.

   \smallskip
   \item $(M, N) \flatresiter N$ if $N$ has no occurrences of $\cst{p}$.
   \end{enumerate}
   \noindent  The \emph{resolved set} $M \flatres_{\!\cst{p}} N$ is the clause set $N'$ such that $(M,N) \flatresiter^* N'$.
   % obtained by the following
   % procedure: Set $N' = N$. Then repeatedly (1)~choose a clause $D \in N'$
   % containing $\cst{p}$, (2)~compute all flat resolvents involving $D$ and
   % partners from $M$, renaming variables apart, and (3)~replace $D$ with these
   % flat resolvents in $N'$. Stop when $N'$ contains no more occurrence of
   % $\cst{p}$.
\end{definitionx}

@@

\begin{definitionx}
   Let $N$ be a clause set and $\cst{p}$ be a singular predicate for
   $N.$ Let $\withpredpos{N}{\cst{p}}$ consist of all clauses of the form $\pospredlit{\cst{p}(\tuple{s}{})} \cor C' \in N$,
   let $\withpredneg{N}{\cst{p}}$ consist of all clauses of the form $\negpredlit{\cst{p}(\tuple{s}{})} \cor C' \in N$,
   let $\withpred{N}{\cst{p}} = \withpredpos{N}{\cst{p}} \cup \withpredneg{N}{\cst{p}}$,
   and let $\withoutpred{N}{\cst{p}} = N \setminus \withpred{N}{\cst{p}}$.
\end{definitionx}
\removemyskip
\begin{definitionx}
   \label{def:pred-elim}
   Let $N$ be a clause set and $\cst{p}$ be a singular predicate for
   $N.$
   \emph{Singular predicate elimination} (SPE) of $\cst{p}$ in $N$ replaces $N$ by
   $\withoutpred{N}{\cst{p}} \cup (\withpredpos{N}{\cst{p}} \flatres_{\!\cst{p}} \withpredneg{N}{\cst{p}})$.
\end{definitionx}

@@

\begin{definitionx}
   \label{def:definition}
   \looseness=-1
   Let $G$ be a clause set, $\cst{p}$ be a predicate symbol, and
   $\tupleempty{x}$ be distinct variables.
%
   The set $G$ is a \emph{definition set} for $\cst{p}$ if (1)~$\cst{p}$
   is singular for $G$, (2)~$G$ consists of clauses of the form
   $\arbpredlit{\cst{p}(\tupleempty{x})} \cor C'$ (up to variable renaming), (3)~the
   variables in $C'$ are all among $\tupleempty{x}$, (4)~all clauses in $\withpredpos{G}{\cst{p}}
   \flatres_{\!\cst{p}} \withpredneg{G}{\cst{p}}$ are tautologies, and
   (5)~$E(\tupleempty{\cst{c}})$ is unsatisfiable, where
   the \emph{environment} $E(\tupleempty{x})$ consists of all subclauses $C'$ of any
   $\arbpredlit{\cst{p}(\tupleempty{x})} \mathbin{\cor} C' \in G$ and
   $\tupleempty{\cst{c}}$ is a tuple of distinct fresh constants substituted
   in for $\tupleempty{x}$.
\end{definitionx}

@@

\begin{examplex}
   \label{example:tautologies}
   Given the formula $\cst{p}(x) \medleftrightarrow \cst{q}(x) \mathrel\land
   (\cst{r}(x) \cor \cst{s}(x))$, a standard clausification algorithm
   \cite{nw-01-small-cnf} produces $\{ \negpredlit{\cst{p}(x)} \cor
   \pospredlit{\cst{q}(x)}{,}\; \negpredlit{\cst{p}(x)} \cor
   \pospredlit{\cst{r}(x)} \cor \pospredlit{\cst{s}(x)}{,}\;
   \pospredlit{\cst{p}(x)} \cor \negpredlit{\cst{q}(x)} \cor
   \negpredlit{\cst{r}(x)}{,}\; \pospredlit{\cst{p}(x)} \cor
   \negpredlit{\cst{q}(x)} \cor \negpredlit{\cst{s}(x)}\}$, which qualifies
   as a definition set for $\cst{p}$.
\end{examplex}

@@

\begin{definitionx}
   \looseness=-1
   Let $N$ be a clause set, $\cst{p}$ be a predicate symbol,
   $G \subseteq N$ be a definition set
   for $\cst{p}$, and $R = \withpred{N}{\cst{p}}
   \setminus G$. \emph{Defined predicate elimination} (DPE) of $\cst{p}$ in $N$ replaces $N$ by
   $\withoutpred{N}{\cst{p}} \cup
%%% @PETAR: Double-check the supression of _p below. --JB
   (\confrep{\withpred{G}{\cst{p}}}{G} \flatres_{\!\cst{p}} \confrep{\withpred{R}{\cst{p}}}{R})$.
\end{definitionx}

@@

   \end{proof}
\end{rep}
\begin{conf}
\removemyskip
\end{conf}
\begin{theoremx}
\label{thm:pes-sat-equiv}
   The result of applying DPE to a clause set $N$ is
   satisfiable if and only if $N$ is satisfiable.
\end{theoremx}
\begin{rep}
\begin{proof}
   Let $\cst{p}$ be a predicate symbol and $G \subseteq N$ be the
   definition set used by DPE, and let $R = \withpred{N}{\cst{p}} \setminus G$.
%
   % First, using Lemmas \ref{lem:flat-res-set-termination-confluence} and \ref{lem:flat-res-set-step-satisfiability}
   % we show that we will eventually obtain a set $R'$
   % that contains no occurrences of $\cst{p}$ and that preserves satisfiability
   % of the original $R$. Finally, we will show that we can
   % omit $G$ in the result without affecting satisfiability.

@@

\begin{definitionx}
   \looseness=-1
   Let $N$ be a clause set and $\cst{p}$ be a predicate symbol. If there exists a
   definition set $G \subseteq N$ for $\cst{p}$, \emph{portfolio predicate
   elimination} (PPE) on $\cst{p}$ in $N$ replaces $N$ with
%%% @PETAR: Double-check the supression of _p below. --JB
   $\withoutpred{N}{\cst{p}} \cup (\confrep{\withpred{G}{\cst{p}}}{G} \flatres_{\!\cst{p}}
   \confrep{\withpred{R}{\cst{p}}}{R})$, where $R = \withpred{N}{\cst{p}} \setminus G$.
   Otherwise, if $\cst{p}$ is singular in $N$, it results in $\withoutpred{N}{\cst{p}}
   \cup (\withpredpos{N}{\cst{p}} \flatres_{\!\cst{p}} \withpredneg{N}{\cst{p}})$.
   In all other cases, it is not applicable.
\end{definitionx}

@@

\begin{examplex}
\label{ex:cex-predelim-sat}
The counterexample below is based on the given clause prover \textsf{GC}
from the saturation framework. It shows how predicate elimination can break
\textsf{GC}'s key invariant, which states that all inferences between active
clauses are redundant. Breaking the invariant means that the limit might be
unsaturated, breaking the refutational completeness proof.

@@

If the prover now eliminates the predicate $\cst{a}$ using \infname{SPE}, the
passive set is reduced to $\{\cst{b} \cor \cst{c} \cor \cst{d}{,}\;
\cst{b} \cor \cst{c} \cor \neglit{\cst{d}}\}$.
%
Either clause is subsumed by an active clause, so the prover deletes it.
It stops with the active set
$\{\cst{c} \cor \cst{d}{,}\; \cst{b} \cor \neglit{\cst{d}}\}$, which is
unsaturated because $\iota$ is no longer redundant. The invariant is broken.
\end{examplex}

\begin{examplex}
\label{ex:cex-predelim-unsat}
In Example~\ref{ex:cex-predelim-sat}, the initial clause set was satisfiable.
If it is unsatisfiable, we can even lose refutational
completeness. To see why, we add the unit clauses
$\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ to the initial clause set of
Example~\ref{ex:cex-predelim-sat} to make it unsatisfiable. We repeat the same
steps as above, including the subsumptions at the end,
yielding the passive set $\{\MAX{\neglit{\cst{b}}}, \MAX{\neglit{\cst{c}}}\}$ and
the active set $\{\cst{c} \cor \MAX{\cst{d}}{,}\; \cst{b} \cor \MAX{\neglit{\cst{d}}}\}$. Then,
making $\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ active triggers no
inferences. The prover stops with an
unsatisfiable four-clause active set that does not contain the empty clause.
\end{examplex}

@@

\begin{definitionx}
\label{def:qle}
Given a clause set, \emph{quasipure literal elimination} (QLE) iteratively
deletes clauses containing quasipure predicate symbols until no such clauses
remain.
\end{definitionx}

>>>>

\newlength{\myskip}
\setlength{\myskip}{0.6ex}
% TYPESETTING: when two environments are next to
% each other, we need to somehow squeeze them
\newcommand{}{\vspace{-\myskip}}

\newenvironment{theorem}% environment name
{% begin code
  \vspace{\myskip}
  \begin{theorem}
}%
{\end{theorem}
 \vspace{\myskip}}% end code

\newenvironment{exa}% environment name
{% begin code
  \vspace{\myskip}
  \begin{example}
}%
{\end{example}
 \vspace{\myskip}}% end code

@@

\newenvironment{defi}% environment name
 {% begin code
   \vspace{\myskip}
   \begin{definition}
 }%
 {\end{definition}
  \vspace{\myskip}}% end code

@@

\begin{defi}
   Given a propositional literal $L$ and a propositional clause set~$N$, the
   set of \emph{propositional hidden literals} for $L$ and~$N$ is
   $\PHL(L, N) = \{ L' \mid L' \bigpimpl{N}^\ast L \} \setminus \{ L \}$, where $\bigpimpl{N}$ is
   defined such that $\neglit{L_1} \bigpimpl{N} L_2$ whenever $L_1 \cor L_2 \in N.$
   Moreover, $\PHL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n
   \PHL(L_i, N)$.
\end{defi}

@@

\begin{defi}
   \label{def:hle}\label{def:hte}
   Given $L' \cor L \cor C \in N,$ if $L' \in
   \PHL(L, N)$, \emph{hidden literal elimination} (HLE) replaces $N$ by $(N
   \setminus \{L' \cor L \cor C\}) \cup \{L \cor C\}$.
   Given $C \in N$, $\{L_1,\dots,L_n\} = \PHL(C, N)$, and $C' = C \cor L_1 \cor
   \cdots \cor L_n$, if $C'$ is a tautology, \emph{hidden
   tautology elimination} (HTE) replaces $N$ by $N \setminus \{C\}$.
\end{defi}

\begin{theorem}
   The result of applying HLE or HTE to a clause set $N$ is
   equivalent to $N$.
\end{theorem}

@@

\begin{defi}
\label{def:hl-fo}
Given a literal $L$ and a clause set $N$,
the set of \emph{hidden literals} for $L$ and $N$
is $\HL(L,N) = \{ L' \mid L' \bigfimpl{N}^\ast L \} \setminus \{ L \}$, where
$\bigfimpl{N}$ is defined so that
(1)~$\neglit{L'\sigma} \bigfimpl{N} L\sigma$ if
$L' \cor L \in N$ and $\sigma$ is a substitution;
(2)~$\eqlit{s}{t} \bigfimpl{N} \eqlit{u[s]}{u[t]}$ for all terms $s, t$ and
contexts $u[\phantom{\cdot}]$; and
(3)~$\neqlit{u[s]}{u[t]} \bigfimpl{N} \neqlit{s}{t}$ for all terms $s, t$ and
contexts $u[\phantom{\cdot}]$.
Moreover, $\HL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n \HL(L_i, N)$.
\end{defi}

@@

\begin{exa}
   \label{example:hle}
   Consider the clause set $N = \{ \cst{p}(x) \cor \neglit{\cst{p}(\cst{f}(x))}{,}\;
   \cst{p}(\cst{f}(\cst{f}(x))) \cor \eqlit{\cst{a}}{\cst{b}} \}$ and the clause
   $C = \neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \cor \cst{p}(x)$. The first
   clause in $N$ induces $\cst{p}(\cst{f}(x)) \bigfimpl{N} {\cst{p}(x)}$,
   $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N} \cst{p}(\cst{f}(x))$, and
   hence $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N}^\ast  \cst{p}(x)$.
   Together with the second clause in $N$, it can be used to derive $\neqlit{\cst{a}}{\cst{b}}
   \bigfimpl{N}^\ast  \cst{p}(x)$. Finally, using rule~(3) of Definition
   \ref{def:hl-fo}, we derive $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})}
   \bigfimpl{N}^\ast  \cst{p}(x)$---that is,
   $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \in \HL(\cst{p}(x), N)$. This
   allows us to remove $C$'s first literal using \infname{HLE}.
\end{exa}

@@

\begin{exa}
   \label{example:unithle}
   \looseness=-1
   Consider the clause set $N = \{\cst{p}(x) \cor \cst{q}(\cst{f}(x)){,}\;
   \neglit{\cst{q}(\cst{f}(\cst{a}))} \cor \eqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}{,}\;
   \neqlit{\cst{f}(x)}{\cst{g}(y)}\}$ and the clause $C = \neglit{\cst{p}(\cst{a})} \cor
   \neglit{\cst{q}(\cst{b})}$. The first clause in $N$ induces
   $\neglit{\cst{q}(\cst{f}(\cst{a}))} \bigfimpl{N} \cst{p}(\cst{a})$, whereas
   the second one induces $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})} \bigfimpl{N}
   \neglit{\cst{q}(\cst{f}(\cst{a}))}$. Thus, we have $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}
   \bigfimpl{N}^\ast \cst{p}(\cst{a})$---that is, $\neqlit{\cst{f}(\cst{b})}{\cst{f}(\cst{c})} \in
   \HL(\cst{p}(\cst{a}), N)$. By applying the substitution $\{x \mapsto \cst{b}{,}\allowbreak\; y
   \mapsto \cst{c}\}$ to the third clause in $N$, we can fulfill the conditions of
   \infname{UnitHLE} and remove $C$'s first literal.
\end{exa}

@@

\begin{defi}
   A clause $C$ is a \emph{hidden tautology} for a clause set $N$ if there
   exists a finite set $\{L_1,\dots,L_n\} \subseteq \HL(C,N)$ such that
   $C \cor L_1 \cor \cdots \cor L_n$ is a tautology.
\end{defi}

\begin{exa}
   \label{example:hidden-tautologies-completeness}
   \looseness=-1
   In general, hidden tautologies are not redundant and cannot be deleted during saturation.
   Consider the unsatisfiable set $N = \{ \negpredlit{\cst{a}}{,}\;
   \negpredlit{\cst{b}}{,}\; \pospredlit{\cst{a}} \cor
   \pospredlit{\cst{c}}{,}\allowbreak\; \pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}
   \}$, the order $\cst{a} \prec
   \cst{b} \prec \cst{c}$, and the empty selection function. The
   only possible superposition inference from $N$ is between the last two clauses,
   yielding the hidden tautology $\pospredlit{\cst{a}} \cor \pospredlit{\cst{b}}$
   (after simplifying away $\neqlit{\top}{\top}$), which is entailed by the larger
   clauses $\pospredlit{\cst{a}} \cor \pospredlit{\cst{c}}$ and
   $\pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}$. If this clause is removed,
   the prover could enter an infinite loop, forever generating and deleting the
   hidden tautology\confrep{}{ and never getting the opportunity to derive the
   empty clause}.
\end{exa}

@@

\begin{theorem}
The rules \infname{HLE}, \infname{FLE}, \infname{CongHLE$^+$},
\infname{Cong\allowbreak HLE$^-$}, \infname{UnitHLE}, \infname{HTR},
\infname{FLR}, and \infname{UnitHTR} are sound simplification rules.
\end{theorem}

@@

\begin{defi}
   A predicate symbol is called \emph{singular}\confrep{ (or ``non-self-referential'')}{}
   for a clause set $N$ if it occurs at most once in every clause contained in
   $N.$
\end{defi}

\begin{defi}
   \label{def:flat-res}
   Let $C = \pospredlit{\cst{p}(\tuplen{s})} \cor C'$ and $D =
   \negpredlit{\cst{p}(\tuplen{t})} \cor D'$ be clauses with no variables in
   common. The clause $\neqlit{s_1}{t_1} \cor \cdots \cor \neqlit{s_n}{t_n}
   \cor C' \cor D'$ is a \emph{flat resolvent} of $C$ and $D$ on $\cst{p}$.
\end{defi}

@@

\begin{defi}
\label{def:flat-res-set}
   Let $M, N$ be clause sets and $\cst{p}$ be a singular predicate for
   $M$. Let $\flatresiter$ be the following relation on clause set pairs and clause sets:
   \begin{enumerate}
   \item $(M{,}\> \{\arbpredlit{\cst{p}(\tupleempty{s})} \cor C'\} \uplus N) \flatresiter (M{,}\; N' \cup N)$ if
   $N'$ is the set that consists of all clauses (up to variable renaming) that are flat resolvents
   with $\arbpredlit{\cst{p}(\tupleempty{s})} \cor C'$ on $\cst{p}$ and a clause from $M$ as premises.
   The premises' variables are renamed apart.

   \smallskip
   \item $(M, N) \flatresiter N$ if $N$ has no occurrences of $\cst{p}$.
   \end{enumerate}
   \noindent  The \emph{resolved set} $M \flatres_{\!\cst{p}} N$ is the clause set $N'$ such that $(M,N) \flatresiter^* N'$.
   % obtained by the following
   % procedure: Set $N' = N$. Then repeatedly (1)~choose a clause $D \in N'$
   % containing $\cst{p}$, (2)~compute all flat resolvents involving $D$ and
   % partners from $M$, renaming variables apart, and (3)~replace $D$ with these
   % flat resolvents in $N'$. Stop when $N'$ contains no more occurrence of
   % $\cst{p}$.
\end{defi}

@@

\begin{defi}
   Let $N$ be a clause set and $\cst{p}$ be a singular predicate for
   $N.$ Let $\withpredpos{N}{\cst{p}}$ consist of all clauses of the form $\pospredlit{\cst{p}(\tuple{s}{})} \cor C' \in N$,
   let $\withpredneg{N}{\cst{p}}$ consist of all clauses of the form $\negpredlit{\cst{p}(\tuple{s}{})} \cor C' \in N$,
   let $\withpred{N}{\cst{p}} = \withpredpos{N}{\cst{p}} \cup \withpredneg{N}{\cst{p}}$,
   and let $\withoutpred{N}{\cst{p}} = N \setminus \withpred{N}{\cst{p}}$.
\end{defi}

\begin{defi}
   \label{def:pred-elim}
   Let $N$ be a clause set and $\cst{p}$ be a singular predicate for
   $N.$
   \emph{Singular predicate elimination} (SPE) of $\cst{p}$ in $N$ replaces $N$ by
   $\withoutpred{N}{\cst{p}} \cup (\withpredpos{N}{\cst{p}} \flatres_{\!\cst{p}} \withpredneg{N}{\cst{p}})$.
\end{defi}

@@

\begin{defi}
   \label{def:definition}
   \looseness=-1
   Let $G$ be a clause set, $\cst{p}$ be a predicate symbol, and
   $\tupleempty{x}$ be distinct variables.
%
   The set $G$ is a \emph{definition set} for $\cst{p}$ if (1)~$\cst{p}$
   is singular for $G$, (2)~$G$ consists of clauses of the form
   $\arbpredlit{\cst{p}(\tupleempty{x})} \cor C'$ (up to variable renaming), (3)~the
   variables in $C'$ are all among $\tupleempty{x}$, (4)~all clauses in $\withpredpos{G}{\cst{p}}
   \flatres_{\!\cst{p}} \withpredneg{G}{\cst{p}}$ are tautologies, and
   (5)~$E(\tupleempty{\cst{c}})$ is unsatisfiable, where
   the \emph{environment} $E(\tupleempty{x})$ consists of all subclauses $C'$ of any
   $\arbpredlit{\cst{p}(\tupleempty{x})} \mathbin{\cor} C' \in G$ and
   $\tupleempty{\cst{c}}$ is a tuple of distinct fresh constants substituted
   in for $\tupleempty{x}$.
\end{defi}

@@

\begin{exa}
   \label{example:tautologies}
   Given the formula $\cst{p}(x) \medleftrightarrow \cst{q}(x) \mathrel\land
   (\cst{r}(x) \cor \cst{s}(x))$, a standard clausification algorithm
   \cite{nw-01-small-cnf} produces $\{ \negpredlit{\cst{p}(x)} \cor
   \pospredlit{\cst{q}(x)}{,}\; \negpredlit{\cst{p}(x)} \cor
   \pospredlit{\cst{r}(x)} \cor \pospredlit{\cst{s}(x)}{,}\;
   \pospredlit{\cst{p}(x)} \cor \negpredlit{\cst{q}(x)} \cor
   \negpredlit{\cst{r}(x)}{,}\; \pospredlit{\cst{p}(x)} \cor
   \negpredlit{\cst{q}(x)} \cor \negpredlit{\cst{s}(x)}\}$, which qualifies
   as a definition set for $\cst{p}$.
\end{exa}

@@

\begin{defi}
   \looseness=-1
   Let $N$ be a clause set, $\cst{p}$ be a predicate symbol,
   $G \subseteq N$ be a definition set
   for $\cst{p}$, and $R = \withpred{N}{\cst{p}}
   \setminus G$. \emph{Defined predicate elimination} (DPE) of $\cst{p}$ in $N$ replaces $N$ by
   $\withoutpred{N}{\cst{p}} \cup
%%% @PETAR: Double-check the supression of _p below. --JB
   (\confrep{\withpred{G}{\cst{p}}}{G} \flatres_{\!\cst{p}} \confrep{\withpred{R}{\cst{p}}}{R})$.
\end{defi}

@@

   \end{proof}
\end{rep}
\begin{conf}

\end{conf}
\begin{theorem}
\label{thm:pes-sat-equiv}
   The result of applying DPE to a clause set $N$ is
   satisfiable if and only if $N$ is satisfiable.
\end{theorem}
\begin{rep}
\begin{proof}
   Let $\cst{p}$ be a predicate symbol and $G \subseteq N$ be the
   definition set used by DPE, and let $R = \withpred{N}{\cst{p}} \setminus G$.
%
   % First, using Lemmas \ref{lem:flat-res-set-termination-confluence} and \ref{lem:flat-res-set-step-satisfiability}
   % we show that we will eventually obtain a set $R'$
   % that contains no occurrences of $\cst{p}$ and that preserves satisfiability
   % of the original $R$. Finally, we will show that we can
   % omit $G$ in the result without affecting satisfiability.

@@

\begin{defi}
   \looseness=-1
   Let $N$ be a clause set and $\cst{p}$ be a predicate symbol. If there exists a
   definition set $G \subseteq N$ for $\cst{p}$, \emph{portfolio predicate
   elimination} (PPE) on $\cst{p}$ in $N$ replaces $N$ with
%%% @PETAR: Double-check the supression of _p below. --JB
   $\withoutpred{N}{\cst{p}} \cup (\confrep{\withpred{G}{\cst{p}}}{G} \flatres_{\!\cst{p}}
   \confrep{\withpred{R}{\cst{p}}}{R})$, where $R = \withpred{N}{\cst{p}} \setminus G$.
   Otherwise, if $\cst{p}$ is singular in $N$, it results in $\withoutpred{N}{\cst{p}}
   \cup (\withpredpos{N}{\cst{p}} \flatres_{\!\cst{p}} \withpredneg{N}{\cst{p}})$.
   In all other cases, it is not applicable.
\end{defi}

@@

\begin{exa}
\label{ex:cex-predelim-sat}
The counterexample below is based on the given clause prover \textsf{GC}
from the saturation framework. It shows how predicate elimination can break
\textsf{GC}'s key invariant, which states that all inferences between active
clauses are redundant. Breaking the invariant means that the limit might be
unsaturated, breaking the refutational completeness proof.

@@

If the prover now eliminates the predicate $\cst{a}$ using \infname{SPE}, the
passive set is reduced to $\{\cst{b} \cor \cst{c} \cor \cst{d}{,}\;
\cst{b} \cor \cst{c} \cor \neglit{\cst{d}}\}$.
%
Either clause is subsumed by an active clause, so the prover deletes it.
It stops with the active set
$\{\cst{c} \cor \cst{d}{,}\; \cst{b} \cor \neglit{\cst{d}}\}$, which is
unsaturated because $\iota$ is no longer redundant. The invariant is broken.
\end{exa}

\begin{exa}
\label{ex:cex-predelim-unsat}
In Example~\ref{ex:cex-predelim-sat}, the initial clause set was satisfiable.
If it is unsatisfiable, we can even lose refutational
completeness. To see why, we add the unit clauses
$\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ to the initial clause set of
Example~\ref{ex:cex-predelim-sat} to make it unsatisfiable. We repeat the same
steps as above, including the subsumptions at the end,
yielding the passive set $\{\MAX{\neglit{\cst{b}}}, \MAX{\neglit{\cst{c}}}\}$ and
the active set $\{\cst{c} \cor \MAX{\cst{d}}{,}\; \cst{b} \cor \MAX{\neglit{\cst{d}}}\}$. Then,
making $\neglit{\cst{b}}$ and $\neglit{\cst{c}}$ active triggers no
inferences. The prover stops with an
unsatisfiable four-clause active set that does not contain the empty clause.
\end{exa}

@@

\begin{defi}
\label{def:qle}
Given a clause set, \emph{quasipure literal elimination} (QLE) iteratively
deletes clauses containing quasipure predicate symbols until no such clauses
remain.
\end{defi}
